%*******************************************************
% Publications
%*******************************************************
\pdfbookmark[1]{Publications}{publications}
\chapter*{Publications}
\section*{First author}
\begin{itemize}
    \item Th\"uring T., Abis M., Wang Z., David C., Stampanoni M.,
        \emph{X-ray phase contrast imaging at 100 keV on a conventional
        source}, Sci. Rep. 4, 2014\footnote{Contributed equally to this work
        with T. Th\"uring.}
    \item Abis M., Lovric G., Schittny J., Stampanoni
    M., \emph{Quantitative lung microstructure mapping with X-ray dark-field
    grating interferometry on a laboratory source}, in preparation
\end{itemize}
\section*{Second author}
\begin{itemize}
    \item Kagias M., Abis M., Wang Z., Jefimovs K., Stampanoni M.,
        \emph{A unfied approach to small angle scattering imaging}, in
        preparation
\end{itemize}
\section*{Contributed author}
\begin{itemize}
    \item Modregger P., Kagias M., Peter S., Abis M., Guzenko V., David C.,
        Stampanoni M., \emph{Multiple scattering tomography}, Phys. Rev.
        Lett. 113, 2014
    \item Gkoumas S., Wang Z., Abis M., Arboleda C., Tudosie G., Donath T.,
        Br\"onnimann C., Schulze-Briese C., Stampanoni M.,
        \emph{Grating-based interferometry and hybrid photon counting
        detectors: Towards a new era in X-ray medical imaging}, Nucl. Inst.
        A 809, 2016
    \item Gkoumas S., Villanueva-Perez P., Wang Z., Romano L., Abis M.,
        Stampanoni M., \emph{A generalized quantitative interpretation of
        dark-field contrast for highly concentrated microsphere
    suspensions}, Sci. Rep. 6, 2016
\end{itemize}
