%********************************************************************
% Appendix
%*******************************************************
% If problems with the headers: get headings in appendix etc. right
%\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
\chapter{Laboratory controls with pyepics and HDF5}
This appendix describes the programs used to control the motors and
detectors used throughout the different experiments.

The documentation refers to version \texttt{495333}, published on 
\href{https://github.com/Enucatl/python-controls-high-energy/tree/495333d062067ada0331b8c688dd4067832981fc}{GitHub}~\parencite{python-controls-high-energy}.

\section{Prerequisites}
The Eiger and Pilatus detector from Dectris Ltd.\ have been used, which
require the \textsc{albula} software~\parencite{albula}. For other detectors and
for the motor controls other packages are required (Python 3.6.4):
\begin{lstlisting}
requests >= 2.18.4
h5py >= 2.7.1
click >= 6.7
numpy >= 1.14.0
scipy >= 1.0.0
decorator >= 4.2.1
pyepics >= 3.3.1
ipython >= 6.2.1
\end{lstlisting}

\section{Installation}
The code is installed on the system's python distribution as a standard
package
\begin{lstlisting}
python setup.py develop
\end{lstlisting}

\section{Motors}
The \texttt{controls.motors.Motor} class controls an EPICS motor, through
its Python interface pyepics. Two methods should be available to the user:
\begin{itemize}
    \item \verb|motor.mv(absolute_position, timeout=9999)| moves the motor
        to the specified absolute position;
    \item \verb|motor.mvr(relative_position, timeout=9999)| moves the motor
        to the specified position, relative to the current position.
\end{itemize}

\section{Detectors}
A generic detector implements the following methods to control the operation
of the detector during scans synchronized with motor movements:
\begin{itemize}
    \item \verb|detector.__init__(host, port, photon_energy, storage_path)|
        connect to the detector at the network address \texttt{host:port}.
        Set an energy threshold \texttt{photon\_energy} (\si{\eV}) and set the
        output directory for the images to \texttt{storage\_path};
    \item \verb|detector.arm()| prepares the detector before an exposure;
    \item \verb|detector.disarm()| hook for post-exposure routines;
    \item \verb|detector.trigger(exposure_time=1)| send a trigger to the
        detector for an exposure of \texttt{exposure\_time} seconds;
    \item \verb|detector.save()| grab all the exposures temporarily stored
        on the detector memory after each \texttt{trigger} call, and save
        them to local destination folder \texttt{storage\_path};
    \item \verb|detector.snap(exposure_time=1)| trigger and save one
        exposure. It calls \texttt{arm}, \texttt{trigger}, \texttt{disarm} and
        \texttt{save}.
\end{itemize}

Calling \texttt{detector.save()} produces a timestamped HDF5 output file inside the
folder \texttt{storage\_path}. The HDF5 file contains all the images
recorded so far, in chronological order, as separate datasets.

\section{Command line operation}
The IPython~\parencite{PER-GRA:2007} package allows for the easy programming of
an enhanced Python interpreter, with the forementioned packages. This
command line is started with the \texttt{bunker4controls} command, with
the following arguments:

\begin{itemize}
    \item the \verb|--storage_path| is used to set the output path for all the images
of the session;
    \item the \verb|--threshold| sets the energy threshold of the detector;
    \item the \verb|-v| flag can be repeated to increase the verbosity
        level;
    \item the \verb|--help| flag prints a description of the available command line
        flags.
\end{itemize}

This command then launches the \texttt{controls/scripts/cli.py} program,
initializing all the motors and detectors

\begin{lstlisting}[language=Python]
import click
import IPython
import logging.config
import logging

import controls.motors
import controls.eiger
import controls.remote_detector
import controls.pilatus
import controls.hamamatsu_flat_panel
import controls.comet_tube
import controls.scans
import controls.log_config

@click.command()
@click.option("-v", "--verbose", count=True)
@click.option("-s", "--storage_path",
    default="/afs/psi.ch/project/hedpc/raw_data/2016/pilatus/2016.07.19",
    type=click.Path(exists=True))
@click.option("-t", "--threshold",
    default=10000,
    help="detector threshold energy (eV)")
def main(verbose, storage_path, threshold):
    logger = logging.getLogger()
    logging.config.dictConfig(controls.log_config.get_dict(verbose))
    g0trx = controls.motors.Motor("X02DA-BNK-HE:G0_TRX", "g0trx")
    g0try = controls.motors.Motor("X02DA-BNK-HE:G0_TRY", "g0try")
    g0trz = controls.motors.Motor("X02DA-BNK-HE:G0_TRZ", "g0trz")
    ...
    stptrx = controls.motors.Motor("X02DA-BNK-HE:STP_TRX", "stptrx")
    detector = controls.eiger.Eiger(
    "129.129.99.112",
    storage_path=storage_path,
    photon_energy=threshold)
    IPython.embed()
\end{lstlisting}

This launches an enhanced IPython command line that can be used to move all
the defined motors and detectors. These definitions should be changed in
\texttt{controls/scripts/cli.py} if necessary.

\section{Scanning procedures}
Scanning procedures are defined in \texttt{controls/scans.py} in order to
coordinate motor movements with detector triggers. The two main functions are
\texttt{dscan} and
\texttt{phase\_stepping\_scan}.

\begin{lstlisting}[language=Python]
def dscan(detector, motor, begin, end, intervals, exposure_time=1)
\end{lstlisting}
this function moves \texttt{motor} from the relative position \texttt{begin}
to \texttt{end} in \texttt{intervals} intervals, calling a
\texttt{detector.trigger(exposure\_time)} at each motor position. It returns
the \texttt{motor} to the original position after the sca to the original
position after the scan.
A single HDF5 output file is saved with one dataset per detector exposure
inside.


\begin{lstlisting}[language=Python]
def phase_stepping_scan(
        detector, motor, begin, end, intervals,
        phase_stepping_motor, phase_stepping_begin, phase_stepping_end,
        phase_steps, exposure_time=1)
\end{lstlisting}
this function moves \texttt{motor} from the relative position \texttt{begin}
to \texttt{end} in \texttt{intervals}, then runs a phase stepping scan at
each \texttt{motor} position. The phase stepping scan runs from
\texttt{phase\_stepping\_begin} to
\texttt{phase\_stepping\_end} (relative), by moving
\texttt{phase\_stepping\_motor} with 
\texttt{phase\_steps}. A \texttt{detector.trigger(exposure\_time)} is called
at each phase step.
All motors are returned to their original positions after the scan.
A single HDF5 output file is saved with one dataset per detector exposure
inside.

\section{X-ray tube}
A module is defined to control the operation of the Comet MXR-225/26 and
MXR-160HP/11 sources. The class \texttt{CometTube} is defined in
\texttt{controls/comet\_tube.py} and communicates over a serial connection
to the tube controller.
The following activities are defined, with serial codes sent to the
controller according to the tube manual\cn.

\begin{itemize}
    \item \verb|CometTube(serial_port="/dev/ttyS0", baudrate=9600, timeout=1)|
        initializes the serial controller on the specified
        serial port, baudrate and communication timeout in seconds;
    \item \verb|current()| reads the current in \si{\milli\ampere}, or
        by passing a value it can set it \verb|current(value)|;
    \item \verb|voltage()| reads the voltage in \si{\kilo\volt}, or
        by passing a value it can set it \verb|voltage(value)|;
    \item \verb|set_small_focus()| and \verb|set_large_focus()| change the
        focal spot size to small and large respectively. These tubes have
        only two focal spot sizes;
    \item \verb|on()| start the X-ray generation. The interface is not able
        to determine if the safety circuit is closed, so you need to check
        that manually;
    \item \verb|off()| turn off the tube.
\end{itemize}
